Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-04-15 Manuscript/manuscriptPSsuppl.tex
Update manuscriptPSsuppl.tex
My edits to the SI NBfix section

\pagebreak
\section{Sensitivity of the molecular electrometer to the chosen definition of ion concentration}
%\section{Effect of the definition of concentration on the headgroup order parameters as a function of ion concentration.}\label{concentrationDEFsection}
Previous studies using the electrometer concept to assess the ion binding affinity to lipid
bilayers report ion concentrations either in water before solvating the lipids (buffer concentration) \cite{akutsu81,roux90,catte16}
or in bulk water after solvating the lipids (bulk concentration) \cite{altenbach84,melcr18}.
In this work, we use the former definition to be consistent with the experimental reference data \cite{roux90}. The difference between these two concentrations increases
with the increasing ion binding affinity. However, Fig. \ref{concentrationDEFfigure} demonstrates that 
in a model with realistic ion binding affinity~\cite{melcr18} the deviation between the two definitions is neglible.
\begin{figure}[]
  \centering
  \includegraphics[width=8.3cm]{../Figs/OP_ECC_POPC_DPPC_water_conc2_dppc_bulk.pdf}
  \caption{\label{concentrationDEFfigure}
    Changes of the headgroup order parameters as a function of CaCl$_2$ concentration using
    two possible definitions of ion concentration from the recent force field
    with realistic calcium binding affinity to a POPC bilayer \cite{melcr18}
    together with the experimental data \cite{akutsu81,altenbach84}. 
  }
\end{figure}

@@

The response of POPC headgroup order parameters to the CaCl$_2$ concentration
are underestimated in simulations of POPC:POPS (5:1) mixture with CHARMM36 containing
the NBfix for interactions between calicum and lipid oxygens \cite{kim16}
(Fig. \ref{changesWITHCaClPS} in the main text),
indicating that the calcium binding to the bilayer is too weak with these
parameters. Without the NBfix term, the calcium binding affinity to pure POPC
lipid bilayers was overestimated in the CHARMM36 model \cite{catte16}.
However, after employing the NBfix term, the response of headgroup order
parameters (Fig. \ref{OP_CHARMM_CaCl_POPC_NBFix}) and the binding affinity
(Fig. \ref{density_profile_CHARMM_CaCl_POPC_NBFix}) of calcium also to a pure
POPC bilayer are underestimated. Notably, CHARMM36 simulations with the NBfix
terms \cite{venable13,kim16} employed predict similar binding affinity for sodium and calcium.
In conclusion, the special NBfix~\cite{kim16} for calcium, 
incorporated in the parameters give by the CHARMM-GUI at the time
of running the simulations (January 2018), leads to the underestimation of
calcium binding affinity to pure PC and mixed PC:PS bilayers.

\begin{figure}[]
  \centering
  \includegraphics[width=17.0cm]{../Figs/OP_CHARMM_CaCl_POPC_NBFix.pdf}
  \caption{\label{OP_CHARMM_CaCl_POPC_NBFix}
    Headgroup order parameters from CHARMM36 simulations of POPC mixed with ions having
    the NBfix term employed for sodium \cite{venable13} ({\it left})
    and calcium \cite{kim16} ({\it right}) compared with the experimental data \cite{akutsu81,altenbach84}
    and simulations without NBfix for the calcium.
    Simulation files without ions are available at Ref. \citenum{nencineCHARMM36popcT310K},
    with the NBfix term in sodium at Ref. \citenum{CHARMM36withCaClNBfix}, 
    with the NBfix term in calcium at Ref. \citenum{CHARMM36withCaClNBfix} and
    without the NBfix in calcium at Ref. \citenum{charmmPOPC450mMCaClfiles}.
  }
\end{figure}

@@

\pagebreak
\section{Calcium density profiles from simulations with \\ POPC:POPS(5:1) mixture}
\begin{figure}[ht]
  \centering
  \includegraphics[width=10cm]{../Figs/CAdensPCPSmixture.eps}
  \caption{\label{CAdensPCPSmixtureALL}
    Ca2+ density profiles from simulations.
  }
  \todo{Should we include also counterions into the plot?} \\
\end{figure}

>>>>

\pagebreak
\section{Effect of the definition of concentration on the headgroup order parameters as a function of ion concentration.}\label{concentrationDEFsection}
Previous studies using the electrometer concept to assess the ion binding affinity to lipid
bilayers report ion concentrations either in water before solvating the lipids (buffer concentration) \cite{akutsu81,roux90,catte16}
or in bulk water after solvating the lipids (bulk concentration) \cite{altenbach84,melcr18}.
In this work, we use the former definition to be consistent with the reference
experimental data \cite{roux90}. The difference between these two concentrations increases
with the increasing ion binding affinity. However, the difference is essentially negligible
in the recent model \cite{melcr18} with realistic ion binding affinity (Fig. \ref{concentrationDEFfigure}).
\begin{figure}[]
  \centering
  \includegraphics[width=8.3cm]{../Figs/OP_ECC_POPC_DPPC_water_conc2_dppc_bulk.pdf}
  \caption{\label{concentrationDEFfigure}
    Changes of the headgroup order parameters as a function of CaCl$_2$ concentration using
    two possible definitions of ion concentration from the recent force field
    with realistic calcium binding affinity to a POPC bilayer \cite{melcr18}
    together with the experimental data \cite{akutsu81,altenbach84}. 
  }
\end{figure}

@@

The response of POPC headgroup order parameters to the CaCl$_2$ concentration
are underestimated in simulations of POPC:POPS (5:1) mixture with CHARMM36 when employing NBfix (as obtained from CHARMM_GUI in January 2018) for interactions between calcium and lipid oxygens \cite{kim16}
(Fig. \ref{changesWITHCaClPS} in the main text),
indicating that the calcium binding to the bilayer is too weak with these
parameters. The response of headgroup order
parameters (Fig. \ref{OP_CHARMM_CaCl_POPC_NBFix}) and the binding affinity
(Fig. \ref{density_profile_CHARMM_CaCl_POPC_NBFix}) of calcium also to a pure
POPC bilayer are underestimated when using NBfix ions.  Notably, CHARMM36 simulations with the NBfix
terms \cite{venable13,kim16} predict similar binding affinity for sodium and calcium. Without the NBfix term, the calcium binding affinity to pure POPC lipid bilayers was overestimated in the CHARMM36 model \cite{catte16}.

%HA: this just repeats the above? 
%In conclusion, the special NBfix~\cite{kim16} for calcium, 
%incorporated in the parameters give by the CHARMM-GUI at the time
%of running the simulations (January 2018), leads to the underestimation of
%calcium binding affinity to pure PC and mixed PC:PS bilayers.

\begin{figure}[]
  \centering
  \includegraphics[width=17.0cm]{../Figs/OP_CHARMM_CaCl_POPC_NBFix.pdf}
  \caption{\label{OP_CHARMM_CaCl_POPC_NBFix}
    Headgroup order parameters from CHARMM36 simulations of POPC, where the NBfix term was employed employed for sodium \cite{venable13} ({\it left})
    and calcium \cite{kim16} ({\it right}) compared with the experimental data \cite{akutsu81,altenbach84}
    and simulations without NBfix for the calcium.
    Simulation files without ions are available at Ref. \citenum{nencineCHARMM36popcT310K},
    with the NBfix term in sodium at Ref. \citenum{CHARMM36withCaClNBfix}, 
    with the NBfix term in calcium at Ref. \citenum{CHARMM36withCaClNBfix} and
    without the NBfix in calcium at Ref. \citenum{charmmPOPC450mMCaClfiles}.
  }
\end{figure}

@@
